---
title: "p8105_hw3_ew2718"
author: "Emma Warshaw"
date: "2022-10-11"
output: github_document
---
```{r}
library(tidyverse)
library(knitr)
```

**Problem 1**
This problem uses the Instacart data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets using:
```{r}
library(p8105.datasets)
data("instacart")
```

#The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

This is a very large dataset: there are 1,384,617 observations for a total of 15 variables.  The variables include: order id, product id, add to cart order, user id, evaluation set, the order number, order_dow, order hour of the day, days since prior order, product name, aisle id, department id, aisle, and department.  Variables are a mix of numberic and character variables.  

#How many aisles are there, and which aisles are the most items ordered from?
There are 134 aisles total.  Aisle 3 has 17,449 observations. 
```{r}
#instacart %>% 
  #group_by(aisle_id) %>%
  #summarize(n_obs = n())
```

#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
#ggplot(instacart, aes(x = aisle_id)) + 
  #geom_histogram() %>%
  #ifelse(instacart >= 10000 )
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

**Problem 2**
This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

#Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

The resulting data frame is 6 columns wide and 50,400 rows long.  There are 6 variables: day_type, week, day_id, day, minute and activity.  For each day_id, there are 1440 rows, corresponding to each minute of the day and the activity score associated with it.  This data is technically tidy, but it's long and not super user friendly.  

```{r}
accelerometer = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               values_to = "activity") %>%
  mutate(day_type = case_when(
    day == 'Monday' ~ 'weekday',
    day == 'Tuesday' ~ 'weekday',
    day == 'Wednesday' ~ 'weekday',
    day == 'Thursday' ~ 'weekday',
    day == 'Friday' ~ 'weekday',
    day == 'Saturday' ~ 'weekend',
    day == 'Sunday' ~ 'weekend'
  )) %>% 
  select(day_type, everything())
```

#Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

The last two Saturdays of the time period (ID=24 & 31) have extremely low activity scores. In fact, the last two weekends have much lower activity scores than the previous two weekends.  

```{r}
accelerometer %>%
  group_by(day_id) %>%
  summarise(activity_total = sum(activity)) %>%
  kable()
```

#Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

Let's start with the obvious: this is a messy graph.  Aside from that, the 14th day has an abnormally high activity score, especially compared to the other Wednesdays in the graph.  The first two weekends of the data set have relatively high activity scores.  The last two weekends, however, drop down to some of the lowest activity scores for the overall time period.  The last two Saturdays are incredibly low, making me wonder if the device was malfunctioning or if the participant removed it for a large protion of the day.    

```{r}
accelerometer %>%
  ggplot(aes(x = day_id, y = activity)) + geom_point(aes(color = day))
```

**Problem 3**
```{r}
library(p8105.datasets)
data("ny_noaa")
```

#The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

There is a problem with missing data.  November of 2007 seems to be missing data for all variables for the 1-16 of the month.  All rows for 2007 are missing temperature minimums and maximums.  Many of the days also look to be missing temperature minimum and maximums.  This is at first glance.  With over 2 million rows worth of information, it's hard to get a sense of the scope of missing data.    

#Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

The most commonly observed value is zero because for most days of the year, it doesn't snow.  

```{r}
ny_noaa_tidy = ny_noaa %>%
  janitor::clean_names() %>%
  separate(col = date, into = c('year', 'month', 'day'), sep = '-') %>%
  mutate(
    month = as.integer(month), 
    month = month.name[month]
  ) %>%
  mutate(
    prcp = prcp*0.01,
    snow = snow*0.1) %>%
  mutate(
    tmax = as.numeric(tmax, na.rm = TRUE), 
    tmin = as.numeric(tmin, na.rm = TRUE)) %>%
  mutate(
    tmax = tmax*0.1,
    tmin = tmin*0.1)

summary_ny = ny_noaa_tidy %>% 
  group_by(snow) %>%
  summarise(n_obs = n())
```

#Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

In January, the average temperature fluctuates around 0 degrees Celsius.  2006 is on average much warmer in January than in other years.  In July, the average temperature each year fluctuates around 25 degrees celsius.  July of 2010 is also warmer on average than the other years in the data set.  Climate change...lovely, isn't it?     

```{r}
ny_noaa_tidy %>% 
  filter(month %in% c("January", "July")) %>% 
  group_by(month, year, id) %>%
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_tmax)) + geom_line(alpha = 0.5, aes(color = year)) + facet_grid(. ~ month)
```

#Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
  ny_noaa_tidy %>%
  ggplot(aes(x = tmin, y = tmax)) + geom_point(alpha = 0.1, aes(color = month), show.legend = FALSE)

  ny_noaa_tidy %>%
  filter(snow > 0, snow <= 100) %>%
  ggplot(aes(x = snow)) + geom_density(aes(color = year))

#plot_1 + plot_2
```





